---
title: "Working Kaggle Notebook"
author: "Roy Emanuel"
output: word_document
---

# Initial Data Plots of the field during a play

The first step I took was to look at the tracking data to see if I
could plot it. This was pretty straightforward. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
library(tidyverse)
library(captioner)
library(pander)
table_nums <- captioner(prefix = "Table")
fig_nums <- captioner()
```

Plot of the first play available. 


```{r echo = FALSE}
library(rstan)
library(rstanarm)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
## load the week one data
w1 <- read_csv("week1.csv")

all_plays <- read_csv("plays.csv") %>%
    filter(!is.na(personnelD))

game_data <- read_csv("games.csv")

w1p1 <-
    w1 %>%
    filter(playId == 75)

## Plot the first pass play
ggplot(w1p1, aes(x, y, color = displayName)) +
    geom_point(alpha = 0.5) +
    xlim(0, 130) +
    ylim(0, 63.3) +
    theme_bw() +
    theme(legend.position = "bottom")
fig_nums(name = "p1", caption = "Movement of all players")
```

Let's look at the players involved with the football: football, Matt
Ryan, Julio Jones and Malcolm Jenkins.

```{r echo = FALSE, fig.width = 7, fig.height = 4}
w1p1ball <-
    w1p1 %>%
    filter(displayName == "Matt Ryan" |
           displayName == "Julio Jones" |
           displayName == "Football" |
           displayName == "Malcolm Jenkins")

ggplot(w1p1ball, aes(x, y, color = displayName)) +
    geom_point(alpha = 0.5) +
    xlim(0, 130) +
    ylim(0, 63.3) +
    theme_bw() +
    theme(legend.position = "bottom")
fig_nums(name = "p1_involved", caption = "Movement of players closest to the football")


ggplot(w1p1ball, aes(time, s, color = displayName)) +
    geom_line() + 
    theme_bw() + 
    theme(legend.position = "bottom")
fig_nums(name = "p1_speed", caption = "Speed of players closest to the football over time")

ggplot(w1p1ball, aes(time, a, color = displayName)) +
    geom_line()  +
    theme(legend.position = "bottom") +
    theme_bw()
fig_nums(name = "p1_speed", caption = "Acceleration of players closest to the football over time")
```

NExt, I'm going to plot a vector for each point showing where the
player is moving and how fast relatively. 

```{r echo=FALSE, fig.width = 7, fig.height = 4}
bld_vector <- function(trkdata) {
    no_football <-
        trkdata %>%
        filter(displayName != "Football")
    fastest <- max(no_football$s)
    trkdata <-
        trkdata %>%
        mutate(sNorm = s / fastest,
               dir_rad = dir * pi / 180,
               o_rad = o * pi / 180, 
               yVel = cos(dir_rad) * sNorm,
               xVel = sin(dir_rad) * sNorm)
    return(trkdata)
}

w1p1 <- bld_vector(w1p1)
ggplot(w1p1, aes(x, y, color = displayName)) +
    geom_point() +
    geom_segment(aes(xend = x + xVel, yend = y + yVel),
                 arrow = arrow(length = unit(0.1, "cm"))) +
    theme_bw() + theme(legend.position = "bottom")
fig_nums(name = "Pos_Speed", caption = "Position plots with speed vector")
```

The above plot doesn't really show much. Of course their direction
aligns with their next point in space. That doesn't do much. So let's
look where everyone is facing. 

The arrow vector shows the direction the player was facing during at
each time step. 

```{r echo=FALSE, fig.width=7, fig.height=4}

ggplot(w1p1, aes(x, y, color = displayName)) +
    geom_point() +
    geom_segment(aes(xend = x + sin(o_rad), yend = y + cos(o_rad) ),
                 arrow = arrow(length = unit(0.1, "cm"))) +
theme_bw() + theme(legend.position = "bottom")
fig_nums(name = "Pos_Orientation", caption = "Position plots with orientation vector")

```


# Passing and closure

Ok, now I'm playing around with the passing a little bit. Let's see
where this goes.

```{r passExplore, echo=FALSE}
releaseSpeed <- w1 %>% group_by(gameId, playId) %>% filter(displayName == "Football") %>% summarise(max_vel = max(s), max_acc = max(a)) %>%
    left_join(., all_plays, by=c("gameId", "playId"))
slowPass <- releaseSpeed %>% filter(max_vel <5)
ggplot(releaseSpeed, aes(max_vel, color = passResult, fill = passResult)) +
    geom_histogram(alpha = 0.5, binwidth = 0.5) +
    theme_bw()
fig_nums(name = "releaseSpeed", caption = "Histogram of Football maximum speeds with outcomes for week 1 2018")

ggplot(releaseSpeed, aes(max_acc, color = passResult, fill = passResult)) +
    geom_histogram(alpha = 0.5, binwidth = 0.5) +
    theme_bw()
fig_nums(name = "releaseAccel", caption = "Histogram of Football maximum accelerations with outcomes for week 1 2018")

```

Comparing the two histograms, I'm surprised to see that speed appears
to be a better way to sort out Sacks from the results than
acceleration. There is a clear difference with over 10 not having any
sacks. Is it important to sort it out? I don't know, but it seemed
important at the time. 


Oh well, that's a bit of a red herring. So what is it we want to do?
Well, let's look at EPA and maybe the personnel. This will be
dependent on a lot of stuff, but here goes

```{r EPA}
epa_pitb <-
    all_plays %>%
    group_by(defendersInTheBox) %>%
    summarise(total_epa = sum(epa))

epa_pO <-
    all_plays %>%
    group_by(personnelO) %>%
    summarise(n_epa = n())

epa_defoff <-
    all_plays %>%
    group_by(defendersInTheBox, personnelO) %>%
    summarise(total_epa = sum(epa)) %>%
    arrange(desc(total_epa))

ggplot(all_plays, aes(y = personnelO, x = epa)) +
    geom_boxplot() +
    theme_bw()
fig_nums(name = "OffPers", caption = "Expected Points Added (epa) for each offensive personnel grouping")

ggplot(all_plays, aes(y = as.character(defendersInTheBox), x = epa)) +
    geom_boxplot() +
    theme_bw()
fig_nums(name = "DefPers", caption = "Expected Points Added (epa) for each defenders in the box grouping")


```

# Analysis

I want to find the relationship between defendersInTheBox, personnelO,
and expected points added (epa). EPA takes down, distance, and field
position into account, so we will consider these issues handled for
now. 

First we will build a linear model. The minimum recorded epa is
-11.94. The maximum recorded epa is 8.629. What is the likelihood?
What is the prior? Let's start with a normal likelihood, normal prior,
and see what comes out. Everything is additive, and away we go...

$$y_{i} \sim \text{Normal}\left(\mu_i, \sigma\right)$$

$$mu_i = \beta_0 + \beta_{def}x_{def,i} + \beta{off}x_{off,i}$$

$$\sigma \sim \text{Gamma}\left(0.1, 0.1\right)$$

```{r firstFit}
all_playsCat <-
    all_plays %>%
    mutate(quarter = as.factor(quarter),
           down = as.factor(down))

## if(file.exists("fitOffDef.Rdata")){
##     load("fitOffDef.Rdata")
## } else {
##     fitOffDef <-
##         stan_glmer(
##             epa ~ defendersInTheBox +
##                 personnelO +
##                 defendersInTheBox +
##                 yardsToGo +
##                 down +
##                 quarter +
##                 1 | personnelD,
##             data = all_plays )
##     save(fitOffDef, file = "fitOffDef.Rdata")
## }


## fitDown <-
##     stan_glm(
##         epa ~ yardsToGo,
##         data = all_plays,
##         family = gaussian("identity"),
##         seed = 12345)

## fitResult <-
##     stan_glm(
##         epa ~ passResult,
##         family = gaussian("identity"),
##         data = all_plays,
##         seed = 12345)
```

Kaggle has a list of questions. Let's start with that:

1. What are coverage schemes that the defense employs? 
1. What coverage options tend to be better performaing?
1. What players are the best at closely tracking receivers as they try
to get open?
1. Which players are the best at closing on receivers when the ball is
   in the air?
1. Which players at defending pass plays when the ball arrives?
1. Is there any way to use player tracking data to predict whether or
   not certain penalties will be called?
1. Who are the NFL's best players against the pass?
1. How does a defense react to certain types of offensive plays?
1. Is there anything about a player that can be used to predict their
   performance on defense?
   
# What are overage schemes that the defense employs

We'll start with the NFL overall, then which teams employ certain
defenses, then which teams have certain defenses employed against
them.

## Overall NFL defensive employment

```{r defEmployment}
## split the personnelD into all types of personnel

all_plays <-
    all_plays %>%
    mutate(DL = str_extract(personnelD, "\\d(?= DL)"),
           LB = str_extract(personnelD, "\\d(?= LB)"),
           DB = str_extract(personnelD, "\\d(?= DB)"),
           WR = str_extract(personnelD, "\\d(?= WR)"),
           TE = str_extract(personnelD, "\\d(?= TE)"),
           RB = str_extract(personnelD, "\\d(?= RB)"))
           

pD <-
    all_plays %>%
    group_by(personnelD) %>%
    summarise(Num = n(),
              Perc = 100 * Num / dim(all_plays)[1]) %>%
    arrange(desc(Num))

pander(pD, caption = table_nums("AllDef", caption = "Types of Defensive Personnel and Frequency for Pass Plays"))

pDL <-
    all_plays %>%
    group_by(DL) %>%
    summarise(Num = n(),
              Perc = 100 * Num / dim(all_plays)[1]) %>%
    arrange(desc(Num))

pLB <-
    all_plays %>%
    group_by(LB) %>%
    summarise(Num = n(),
              Perc = 100 * Num / dim(all_plays)[1]) %>%
    arrange(desc(Num))

pDB <-
    all_plays %>%
    group_by(DB) %>%
    summarise(Num = n(),
              Perc = 100 * Num / dim(all_plays)[1]) %>%
    arrange(desc(Num))

pander(pDL, caption = table_nums("DLDef", caption = "Number of Defensive Linemen and Frequency for Pass Plays"))
pander(pLB, caption = table_nums("LBDef", caption = "Number of Linebackers and Frequency for Pass Plays"))
pander(pDB, caption = table_nums("DBDef", caption = "Number of Defensive Backs and Frequency for Pass Plays"))

```

Let's see how the composition of the linemen affect the outcome of the play.

```{r trimDef}

pOutcomeDL <-
    all_plays %>%
    group_by(DL, passResult) %>%
    summarise(Num = n()) %>%
    arrange(desc(DL))

totOutcome <-
    pOutcomeDL %>%
    summarise(typeSum = sum(Num))

pOutcomeDL <-
    left_join(pOutcomeDL, totOutcome, by = "DL") %>%
    mutate(percentage = 100 *  Num / typeSum)

pOutcomeLB <-
    all_plays %>%
    group_by(LB, passResult) %>%
    summarise(Num = n()) %>%
    arrange(desc(LB))

totOutcome <-
    pOutcomeLB %>%
    summarise(typeSum = sum(Num))

pOutcomeLB <-
    left_join(pOutcomeLB, totOutcome, by = "LB") %>%
    mutate(percentage = 100 *  Num / typeSum)

pOutcomeDB <-
    all_plays %>%
    group_by(DB, passResult) %>%
    summarise(Num = n()) %>%
    arrange(desc(Num))

totOutcome <-
    pOutcomeDB %>%
    summarise(typeSum = sum(Num))

pOutcomeDB <-
    left_join(pOutcomeDB, totOutcome, by = "DB") %>%
    mutate(percentage = 100 *  Num / typeSum)

pOutcomeDL %>%
    arrange(DL) %>%
    pander(caption = "Pass Outcomes by Number of Defensive Linemen")

pOutcomeLB %>%
    arrange(LB) %>%
    pander(caption = "Pass Outcomes by Number of Linebackers")

pOutcomeDB %>%
    arrange(DB) %>%
    pander(caption = "Pass Outcomes by Number of Defensive Backs")

```
